taxize - taxonomic search and retrieval in R

### 2013-07-11 inst/doc/taxize_withcode_plos.Rnw
fixed matching example

\subsection*{Matching species tables with different taxonomic resolution}
Biologist often need to match different sets of data tied to species. Trait-based approaches are a promising tool in ecology \cite{statzner_can_2010}. One problem is that abundance data must be matched with trait databases like \cite{usseglio-polatera_biological_2000}. These two data tables may contain species information on different taxonomic levels and perhabs data must be aggregated to a joint taxomic level, so that the data can be merged. taxize can help in this data-cleaning step, providing a reproducible workflow.

<<tax_match, cache=TRUE>>=
"<TODO>"
@

A more complicated example is given in Appendix B.

>>>>

\subsection*{Matching species tables with different taxonomic resolution}
Biologist often need to match different sets of data tied to species. For example, trait-based approaches are a promising tool in ecology \cite{statzner_can_2010}. One problem is that abundance data must be matched with trait databases like \cite{usseglio-polatera_biological_2000}. These two data tables may contain species information on different taxonomic levels and possibly data must be aggregated to a joint taxomic level, so that the data can be merged. taxize can help in this data-cleaning step, providing a reproducible workflow:

We can use the mentioned \emph{classification}-function to retrieve the taxonomic hierarchy and then search the hierarchies up- and downwards for matches. Here is an example to match a species with names on three different taxonomic levels.

<<tax_match, cache=TRUE, tidy=FALSE>>=
A <- "gammarus roeseli"

B1 <- "gammarus roeseli"
B2 <- "gammarus"
B3 <- "gammaridae"

A_clas <- classification(A, db = 'ncbi')
B1_clas <- classification(B1, db = 'ncbi')
B2_clas <- classification(B2, db = 'ncbi')
B3_clas <- classification(B3, db = 'ncbi')

B1[match(A, B1)]
A_clas[[1]]$Rank[tolower(A_clas[[1]]$ScientificName) %in% B2]
A_clas[[1]]$Rank[tolower(A_clas[[1]]$ScientificName) %in% B3]
@

If we find a direct match (here \emph{Gammarus roeseli}), we are lucky. But we can also match Gammaridae with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait-table to a abundance table) is given in Appendix B.

### 2013-07-11 inst/doc/taxize_withcode_plos.tex
fixed matching example

\subsection*{Matching species tables with different taxonomic resolution}
Biologist often need to match different sets of data tied to species. Trait-based approaches are a promising tool in ecology \cite{statzner_can_2010}. One problem is that abundance data must be matched with trait databases like \cite{usseglio-polatera_biological_2000}. These two data tables may contain species information on different taxonomic levels and perhabs data must be aggregated to a joint taxomic level, so that the data can be merged. taxize can help in this data-cleaning step, providing a reproducible workflow.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstring{"<TODO>"}
\end{alltt}
\begin{verbatim}
[1] "<TODO>"
\end{verbatim}
\end{kframe}
\end{knitrout}

A more complicated example is given in Appendix B.

>>>>

\subsection*{Matching species tables with different taxonomic resolution}
Biologist often need to match different sets of data tied to species. For example, trait-based approaches are a promising tool in ecology \cite{statzner_can_2010}. One problem is that abundance data must be matched with trait databases like \cite{usseglio-polatera_biological_2000}. These two data tables may contain species information on different taxonomic levels and possibly data must be aggregated to a joint taxomic level, so that the data can be merged. taxize can help in this data-cleaning step, providing a reproducible workflow:

We can use the mentioned \emph{classification}-function to retrieve the taxonomic hierarchy and then search the hierarchies up- and downwards for matches. Here is an example to match a species with names on three different taxonomic levels.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
A <- \hlstring{"gammarus roeseli"}

B1 <- \hlstring{"gammarus roeseli"}
B2 <- \hlstring{"gammarus"}
B3 <- \hlstring{"gammaridae"}

A_clas <- \hlfunctioncall{classification}(A, db = \hlstring{'ncbi'})
B1_clas <- \hlfunctioncall{classification}(B1, db = \hlstring{'ncbi'})
B2_clas <- \hlfunctioncall{classification}(B2, db = \hlstring{'ncbi'})
B3_clas <- \hlfunctioncall{classification}(B3, db = \hlstring{'ncbi'})

B1[\hlfunctioncall{match}(A, B1)]
\end{alltt}
\begin{verbatim}
[1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B2]
\end{alltt}
\begin{verbatim}
[1] "genus"
\end{verbatim}
\begin{alltt}
A_clas[[1]]$Rank[\hlfunctioncall{tolower}(A_clas[[1]]$ScientificName) %in% B3]
\end{alltt}
\begin{verbatim}
[1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

If we find a direct match (here \emph{Gammarus roeseli}), we are lucky. But we can also match Gammaridae with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait-table to a abundance table) is given in Appendix B.
